The Proofs

### 2024-07-16 proofs/tex/singularity-lemmas.tex
WIP: fix references, add appendix

            Notice $[[Γ ⊢ iP' : scE]]$ means $[[Γ ⊢ iP' ≥ ∃nbs.↓iN1]]$.
            Let us show that it implies $[[Γ ⊢ iP' ≈ ∃α⁻.↓α⁻]]$.

            $
            \begin{aligned}
            [[Γ ⊢ iP' ≥ ∃nbs.↓iN1]] &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(iN1)]]\\
                                    &\phantom{\Rightarrow} \text{(where $[[ord {nbs} in iN' = nbs']]$)}
                                    && \text{by \cref{corollary:nf-pres-subt}} \\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(nbj)]]  
                                    && \text{by \cref{lemma:normalization-completeness}}\\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nbn]]  
                                    && \text{by definition of normalization}\\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbj.↓nbj]]  
                                    && \text{since $[[ord {nbs} in nf(iN1)]] = [[nbj]]$}\\
                                    &\Rightarrow [[Γ, nbj ⊢ nf(iP') ≥ ↓nbj]] \\ 
                                    &\phantom{\Rightarrow} \text { and } [[nbj ∉ fv(nf(iP'))]]
                                    && \text{by \cref{lemma:shape-supertypes-norm}}\\
            \end{aligned}
            $

            By \cref{lemma:shape-supertypes-norm}, 
            the last subtyping means that $[[nf(iP') = ∃nas.↓iN']]$,
            such that
            \begin{enumerate}
                \item $[[Γ, nbj, nas ⊢ iN']]$
                \item $[[ord {nas} in iN' = nas]]$
                \item for some substitution $[[Γ, nbj ⊢ σ :{nas}]]$, 
                    $[[ [σ]iN' = nbj ]]$.
            \end{enumerate}
            Since $[[nbj ∉ fv(nf(iP'))]]$,
            the latter means that $[[iN' = na]]$, and then 
            $[[nf(iP') = ∃na.↓na]]$ for some $[[na]]$.
            Finally, notice that all the types of shape
            $[[∃na.↓na]]$ are equal.
   \end{caseof}

>>>>

            Notice $[[Γ ⊢ iP' : scE]]$ means $[[Γ ⊢ iP' ≥ ∃nbs.↓iN1]]$.
            Let us show that it implies $[[Γ ⊢ iP' ≈ ∃α⁻.↓α⁻]]$.
            \begin{align*}
            [[Γ ⊢ iP' ≥ ∃nbs.↓iN1]] &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(iN1)]]\\
                                    &\phantom{\Rightarrow} \text{(where $[[ord {nbs} in iN' = nbs']]$)}
                                    && \text{by \cref{corollary:nf-pres-subt}} \\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nf(nbj)]]  
                                    && \text{by \cref{lemma:normalization-completeness}}\\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbs'.↓nbn]]  
                                    && \text{by definition of normalization}\\
                                    &\Rightarrow [[Γ ⊢ nf(iP') ≥ ∃nbj.↓nbj]]  
                                    && \text{since $[[ord {nbs} in nf(iN1)]] = [[nbj]]$}\\
                                    &\Rightarrow [[Γ, nbj ⊢ nf(iP') ≥ ↓nbj]] \\ 
                                    &\phantom{\Rightarrow} \text { and } [[nbj ∉ fv(nf(iP'))]]
                                    && \text{by \cref{lemma:shape-supertypes-norm}}\\
            \end{align*}
            By \cref{lemma:shape-supertypes-norm}, 
            the last subtyping means that $[[nf(iP') = ∃nas.↓iN']]$,
            such that
            \begin{enumerate}
                \item $[[Γ, nbj, nas ⊢ iN']]$
                \item $[[ord {nas} in iN' = nas]]$
                \item for some substitution $[[Γ, nbj ⊢ σ :{nas}]]$, 
                    $[[ [σ]iN' = nbj ]]$.
            \end{enumerate}
            Since $[[nbj ∉ fv(nf(iP'))]]$,
            the latter means that $[[iN' = na]]$, and then 
            $[[nf(iP') = ∃na.↓na]]$ for some $[[na]]$.
            Finally, notice that all the types of shape
            $[[∃na.↓na]]$ are equal.
   \end{caseof}

### 2024-07-16 proofs/tex/translation-proofs.tex
WIP: fix references, add appendix

\lemmaTypPolWf*
\begin{proof}
  We prove it by induction on $[[fΓ ⊢ fT]]$.
  Let us consider the last rule applied to infer this judgment.
  \begin{caseof}
    \item \ruleref{\ottdruleSFwfVarLabel}
      By inversion, $[[a ∊ fΓ]]$, 
      which implies $[[+|a|+]] = [[α⁺ ∊ +|fΓ|+]]$.
      Then $[[+|fΓ|+ ⊢ +|a|+ ]]$
      by \ruleref{\ottdruleWFTPVarLabel};
    \item \ruleref{\ottdruleSFwfArrowLabel}
      By inversion,  $[[fΓ ⊢ fA]]$ and $[[fΓ ⊢ fB]]$,
      which by the induction hypothesis implies $[[+|fΓ|+ ⊢ +|fA|+]]$ and $[[+|fΓ|+ ⊢ +|fB|+]]$.
      Then $[[+|fΓ|+ ⊢ ↓(+|fA|+ → ↑+|fB|+)]] = [[+|fA → fB|+]]$ by 
      \ruleref{\ottdruleWFTShiftDLabel},
      \ruleref{\ottdruleWFTShiftULabel}, and
      \ruleref{\ottdruleWFTArrowLabel};
    \item \ruleref{\ottdruleSFwfForallLabel}
      By inversion, $[[fΓ, a ⊢ fT]]$,
      which by the induction hypothesis implies $[[+|fΓ|+, α⁺ ⊢ +|fT|+]]$.
      Then $[[+|fΓ|+ ⊢ ∀α⁺.↑+|fT|+]] = [[+|∀a.fT|+]]$ by 
      \ruleref{\ottdruleWFTForallLabel}, and \ruleref{\ottdruleWFTShiftULabel}.
  \end{caseof}
\end{proof}

>>>>

\lemmaTypPolWf*
\begin{proof}
  We prove it by induction on $[[fΓ ⊢ fT]]$.
  Let us consider the last rule applied to infer this judgment,
  or, equivalently, the shape of $[[fΓ ⊢ fT]]$.
  \begin{caseof}
    \item $[[fΓ ⊢ a]]$.
      By inversion, $[[a ∊ fΓ]]$, 
      which implies $[[+|a|+]] = [[α⁺ ∊ +|fΓ|+]]$.
      Then $[[+|fΓ|+ ⊢ +|a|+ ]]$
      by \ruleref{\ottdruleWFTPVarLabel};
    \item $[[fΓ ⊢ fA → fB]]$.
      By inversion,  $[[fΓ ⊢ fA]]$ and $[[fΓ ⊢ fB]]$,
      which by the induction hypothesis implies $[[+|fΓ|+ ⊢ +|fA|+]]$ and $[[+|fΓ|+ ⊢ +|fB|+]]$.
      Then $[[+|fΓ|+ ⊢ ↓(+|fA|+ → ↑+|fB|+)]] = [[+|fA → fB|+]]$ by 
      \ruleref{\ottdruleWFTShiftDLabel},
      \ruleref{\ottdruleWFTShiftULabel}, and
      \ruleref{\ottdruleWFTArrowLabel};
    \item $[[fΓ ⊢ ∀as.fT]]$.
      By inversion, $[[fΓ, a ⊢ fT]]$,
      which by the induction hypothesis implies $[[+|fΓ|+, α⁺ ⊢ +|fT|+]]$.
      Then $[[+|fΓ|+ ⊢ ∀α⁺.↑+|fT|+]] = [[+|∀a.fT|+]]$ by 
      \ruleref{\ottdruleWFTForallLabel}, and \ruleref{\ottdruleWFTShiftULabel}.
  \end{caseof}
\end{proof}

### 2024-07-16 proofs/tex/upgrade-lemmas.tex
WIP: fix references, add appendix

\lemmaUpgradeSoundness*
\begin{lemma}[Soundness of Upgrade]\label{lemma:upgrade-soundness}
    Assuming $[[iP]]$ is well-formed in $[[Γ = Δ, pnas]]$,
    if $[[upgrade Γ ⊢ iP to Δ = iQ]]$
    then
    \begin{enumerate}
        \item $[[Δ ⊢ iQ]]$
        \item $[[Γ ⊢ iQ ≥ iP]]$
    \end{enumerate}
\end{lemma}
\begin{proof}
    By inversion, $[[upgrade Γ ⊢ iP to Δ = iQ]]$ means that 
    for fresh $[[pnbs]]$ and $[[pncs]]$,
    $[[Δ, pnbs, pncs ⊨ [pnbs/pnas]iP ∨ [pncs/pnas]iP = iQ]]$.
    Then by the soundness of the least upper bound (\cref{lemma:lub-soundness}),
    \begin{enumerate}
        \item $[[Δ, pnbs, pncs ⊢ iQ]]$, 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$, and 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pncs/pnas]iP]]$.
    \end{enumerate}

    $ 
    \begin{aligned}
        [[fv iQ]] &\subseteq [[fv [pnbs/pnas]iP ∩ fv [pncs/pnas]iP]]
                  &&\text{since by \cref{lemma:fv-propagation}, 
                         $[[fv iQ ⊆ fv [pnbs/pnas]iP]]$ and
                         $[[fv iQ ⊆ fv [pncs/pnas]iP]]$}\\
                  &\subseteq [[ ((fv iP \ {pnas}) ∪ {pnbs}) ∩ ((fv iP \ {pnas}) ∪ {pncs})]]\\
                  &= [[ (fv iP \ {pnas}) ∩ (fv iP \ {pnas}) ]]
                  &&\text{since $[[pnbs]]$ and $[[pncs]]$ are fresh}\\
                  &= [[ fv iP \ {pnas} ]]\\
                  &\subseteq [[ Γ \ {pnas} ]]
                  &&\text{since $[[iP]]$ is well-formed in $[[Γ]]$}\\
                  &\subseteq [[ {Δ} ]]\\
    \end{aligned}
    $\\
    This way, by \cref{lemma:wf-ctxt-equiv}, $[[Δ ⊢ iQ]]$.

>>>>

\lemmaUpgradeSoundness*
\begin{lemma}[Soundness of Upgrade]\label{lemma:upgrade-soundness}
    Assuming $[[iP]]$ is well-formed in $[[Γ = Δ, pnas]]$,
    if $[[upgrade Γ ⊢ iP to Δ = iQ]]$
    then
    \begin{enumerate}
        \item $[[Δ ⊢ iQ]]$
        \item $[[Γ ⊢ iQ ≥ iP]]$
    \end{enumerate}
\end{lemma}
\begin{proof}
    By inversion, $[[upgrade Γ ⊢ iP to Δ = iQ]]$ means that 
    for fresh $[[pnbs]]$ and $[[pncs]]$,
    $[[Δ, pnbs, pncs ⊨ [pnbs/pnas]iP ∨ [pncs/pnas]iP = iQ]]$.
    Then by the soundness of the least upper bound (\cref{lemma:lub-soundness}),
    \begin{enumerate}
        \item $[[Δ, pnbs, pncs ⊢ iQ]]$, 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$, and 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pncs/pnas]iP]]$.
    \end{enumerate}
    \begin{align*}
        [[fv iQ]] &\subseteq [[fv [pnbs/pnas]iP ∩ fv [pncs/pnas]iP]]
                  &&\text{since by \cref{lemma:fv-propagation}, 
                         $[[fv iQ ⊆ fv [pnbs/pnas]iP]]$}\\
                  & &&  \text{and $[[fv iQ ⊆ fv [pncs/pnas]iP]]$}\\
                  &\subseteq [[ ((fv iP \ {pnas}) ∪ {pnbs}) ∩ ((fv iP \ {pnas}) ∪ {pncs})]]\\
                  &= [[ (fv iP \ {pnas}) ∩ (fv iP \ {pnas}) ]]
                  &&\text{since $[[pnbs]]$ and $[[pncs]]$ are fresh}\\
                  &= [[ fv iP \ {pnas} ]]\\
                  &\subseteq [[ Γ \ {pnas} ]]
                  &&\text{since $[[iP]]$ is well-formed in $[[Γ]]$}\\
                  &\subseteq [[ {Δ} ]]\\
    \end{align*}
    This way, by \cref{lemma:wf-ctxt-equiv}, $[[Δ ⊢ iQ]]$.
